---
title: "Rで学ぶ『完全独習 統計学入門』"
author: "Sampo Suzuki, CC 4.0 BY-NC-SA"
date: "`r format(Sys.Date(), format = '%Y-%m-%d')`"
output: pdf_document
description: 本書はデータ分析勉強会で学んだ成果をまとめたものです
site: bookdown::bookdown_site
github-repo: k-metrics/bookdown
always_allow_html: yes
bibliography:
- bib/references.bib
- bib/packages.bib
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fig_crop: no
fontsize: 12pt
linestretch: 1.1
links-as-notes: yes
mainfontoptions: Scale=MatchUppercase
monofontoptions: Scale=0.8
CJKoptions: Scale=1.0
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,                 # Rコードを出力する
                      warning = FALSE,             # Rの警告を抑止する
                      fig_caption = TRUE,
                      fig.align = 'center',        # 図を中央揃えにする
                      attr.source='.numberLines',  # Rコードに行番号表示
                      out.width = "90%"            # 画像幅のデフォルト
                      )

# 出力がPDF(LaTeX)の場合の追加設定
if (knitr::is_latex_output()) {
    knitr::opts_chunk$set(dev = "cairo_pdf",       # 図の出力にPDFデバイスを使う
                          # dev.args = list(family = "Noto Serif"),
                          dev.args = list(family = "Source Han Code JP"),
                                                   # 図のフォントを指定する
                          fig.pos = "H",           # 図をその場で表示させる
                          out.extra = "",
                          tinytex.verbose = TRUE   # デバッグ用メッセージを出力
                          )
# ePUB形式の場合の追加設定
} else if (!knitr::is_html_output(excludes = "epub")) {
    knitr::opts_chunk$set(dev = "jpeg")            # 図の出力にJPEGデバイスを使う
}

# 図の描画位置調整については以下を参照
# https://gedevan-aleksizde.github.io/rmarkdown-cookbook/figure-placement.html#force-floats-forward
# opts_chunkで"fig.pos"を指定することで全てのチャンク出力に対して調整可能
# h: フロートをここ (here) に配置します. つまりソーステキスト上に現れるところとほぼ同じ位置です.
# t: そのページの先頭 (top) に配置します.
# b: そのページの末尾 (bottom) に配置します.
# p: フロート専用の特別なページに配置します.
# !: LaTex が「良い」フロートの位置を決定するための内部パラメータ上書きします.
# H: フロートを正確に LaTex コード上と同じ位置に配置します. float パッケージが必要です (\usepackage{float}).

htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# Update packages.bib file
tidyverse::tidyverse_packages() %>% 
  c(., tidymodels::tidymodels_packages(), "knitr", "rmarkdown","bookdown",
    "DBI", "zoo", "htmlwidgets", "shiny", "modeest", "patchwork", "Rcmdr",
    "e1071", "statip") %>% 
  sort() %>% unique() %>% 
  knitr::write_bib("./bib/packages.bib")

## 出力によりテーブル表記を変更する
# HTML format use rmarkdown::paged_table
# PDF/DOCX format use kableExtra package
#  https://haozhu233.github.io/kableExtra/
#  https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf
# WARNING:
# Should NOT exec df_print befor skimr::skim (has bug) 
# Do NOT require() or library() kableExtra package with using skimr package
df_print <- function(df = NULL,             # 表示させたいデータをフレームを指定
                     n = 3L,                # 省略表示時の行数
                     caption = NULL,        # 表見出しを指定
                     booktabs = TRUE,       # FALSEで縦罫線を描画
                     font_size = NULL,      # 表内文字のフォントサイズ
                     position = "center",   # 表の横位置
                     full_width = FALSE,    # TRUEで幅いっぱいに拡大表示
                     scale_down = FALSE,    # TRUEで横幅に収まるよう縮小表示
                     all = FALSE,           # TRUEですべてのデータを表示
                     head_tail = FALSE,     # TRUEで先頭・末尾をn行ずつ表示
                     ...) {                 # kableExtra::kblに渡す引数
  if (!is.null(df)) {
    if (knitr::is_latex_output()) {
      if (scale_down == FALSE) {
        # latex_option = c("striped", "HOLD_position")
        latex_option = c("striped", "HOLD_position")
      } else {
        # latex_option = c("striped", "HOLD_position", "scale_down")
        latex_option = c("striped", "HOLD_position", "scale_down")
      }
      if (all == FALSE) {
        df <- df %>% 
          psych::headTail(top = n,
                          bottom = dplyr::if_else(head_tail == FALSE, 0L, n))
      }
      df %>% 
        kableExtra::kbl(caption  = caption, booktabs = booktabs, ...) %>%
        kableExtra::kable_styling(font_size = font_size, position = position,
                                  full_width = full_width,
                                  latex_options = latex_option)
    } else {   # 出力タイプ不明の場合
      df %>% 
        rmarkdown::paged_table()
    }
  }
}
```

# Introduction

　本資料は[『完全独習 統計学入門』小島寛之著](https://www.diamond.co.jp/book/9784478820094.html)を用いてRを学ぶために作成しています。 　

　[『完全独習 統計学入門』小島寛之著](https://www.diamond.co.jp/book/9784478820094.html)に記載されているデータの一部をCSVファイルにして利用していますが、データ自体の著作権は原著作者にあります。また、CSVファイルは著作権を侵害する意図で作成したものではありません。  

<!--chapter:end:index.Rmd-->

---
title: "データ分析勉強会2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## スケージュールなど
　詳細は[勉強会ページ](https://sites.google.com/view/kanto-metrics/2022%E5%B9%B4%E5%BA%A6)にてご確認ください。参加者募集中。

## 本年度テキスト
　[『完全独習 統計学入門』小島寛之著](https://www.diamond.co.jp/book/9784478820094.html)

### サンプルデータに関して
　本リポジトリには本年度テキストに記載されているデータの一部をCSVファイルにしていますが、データ自体の著作権は原著作者にあります。  
　CSVファイルは私的利用のために作成したものでありデータの正確性・整合性については保証していません。また、著作権を侵害する意図で作成したものではありません。  
　なお、本年度テキストの第一章に掲載されているデータの原データに関しては[こちらのサイト](http://goldenstate.cocolog-nifty.com/blog/2015/02/111pp2-8-19e6.html)が参考になりますが一部、異なったデータがあるようです。


---
[CC 4.0 BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/deed.ja), Sampo Suzuki (Update: `r lubridate::now()`)

<!--chapter:end:README.Rmd-->

---
title: "Rで学ぶ『完全独習 統計学入門』"
author: "Sampo Suzuki, CC 4.0 BY-NC-SA"
date: "`r format(Sys.Date(), format = '%Y-%m-%d')`"
output: pdf_document
description: 本書はデータ分析勉強会で学んだ成果をまとめたものです
site: bookdown::bookdown_site
github-repo: k-metrics/bookdown
always_allow_html: yes
bibliography:
- bib/references.bib
- bib/packages.bib
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fig_crop: no
fontsize: 12pt
linestretch: 1.1
links-as-notes: yes
mainfontoptions: Scale=MatchUppercase
monofontoptions: Scale=0.8
CJKoptions: Scale=1.0
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(tufte)
require(tidyverse)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

# 第1章の原データ
histgram_data <- "http://goldenstate.cocolog-nifty.com/blog/files/data.txt"
```

# Introduction

　本資料は　[『完全独習 統計学入門』小島寛之著](https://www.diamond.co.jp/book/9784478820094.html)を用いてRを学ぶために作成しています。 　

```{r, include=FALSE}
# df <- histgram_data %>% 
#   readr::read_csv() %>% 
#   dplyr::slice_head(n = 80) %>% 
#   dplyr::select(height, weight)

```

# 第1章

```{r, include=FALSE}
x <- "data/P16_図表1-1 .csv" %>% 
  readr::read_csv(col_names = FALSE) %>% 
  tidyr::pivot_longer(cols = starts_with("X")) %>% 
  dplyr::arrange(name) %>% 
  dplyr::select(height = value) %>% 
  dplyr::mutate(height = as.integer(height))
```

　対象となるデータはP16 図表1-1で示されているデータですが、Rで処理しやすいように以下のように一列にまとめて`x`というデータフレーム型の変数に格納しておきます。

```{r}
x
```

## ヒストグラムを作る
　Rでヒストグラムを描くには`hist()`関数を使うことで描けます。同時に階級、度数と階級値などを求めることもできます。
```{r}
hist <- hist(x$height, include.lowest = FALSE, plot = TRUE)
hist
```

　度数分布表として表にしたい場合は下記のように
```{r}
x %>% 
  # 階級を求めて、データを階級ごとに分ける
  dplyr::mutate(class = cut(height,
                            breaks = pretty(height, n = nclass.Sturges(height)),
                            include.lowest = FALSE, right = TRUE)) %>% 
  # 階級ごとの度数を求める
  dplyr::count(class) %>% 
  # 累積度数、相対度数、累積相対度数を求める
  dplyr::mutate(cumsum_n = cumsum(n),
                prop = prop.table(n), cumsum_prop = cumsum(prop)) %>% 
  # 階級値を求める
  dplyr::mutate(class_value = as.character(class)) %>%
  tidyr::separate(class_value, into = c("l", "h"), sep = ",") %>% 
  dplyr::mutate(l = as.integer(stringr::str_remove(l, "[[:punct:]]")) + 1L,
                h = as.integer(stringr::str_remove(h, "[[:punct:]]")),
                mids = (l + h) / 2L) %>%
  # 変数名を日本語にする
  dplyr::select(`階級` = class, `階級値` = mids,
                `度数` = n, `累積度数` = cumsum_n,
                `相対度数` = prop, `累積相対度数` = cumsum_prop)
```

### 階級を求める
```{r}
with(x, pretty(height, n = nclass.Sturges(height)))
```

　`pretty()`関数は`x`引数で指定するデータから階級を求める関数です。`n`引数で階級を区切る数（階級数）を指定します。ここでは、階級数`n`はスタージェスの公式から求めています。指定した階級数でデータの範囲を単純に割っている訳ではないことに注意してください。`x`で指定したデータが階級に収まるように適切な丸め処理を行いますので、指定した階級数どおりの階級が求められるとは限りません。  

　スタージェスの公式とは、階級数を$k$、データのサイズ（数）を$n$とした場合、下式で定義されます。

$$k = \log_2(n) + 1$$

　Rでは`nclass.Sturgess()`関数が用意されています。
```{r}
with(x, nclass.Sturges(height))
```
　
　スタージェスの公式以外にもスコットの選択（`nclass.scott()`）やフリードマン=ダイアコニスの選択（`nclass.FD()`）などがあります。


### データを各階級に分類する
```{r}
with(x, cut(height, breaks = pretty(height, n = nclass.Sturges(height))))
```

　`cut()`関数はデータがどの階級に入るかを計算する関数です。`breaks`引数に任意の階級を指定することも可能です。

#### 階級の境界値はどちらに含まれるのか？
　`cut()`関数の出力は、`(140,145] (145,150]`のように階級間で同じ数値が含まれます。`145`はどちらの階級に含まれるかは、添えられている括弧で区別できるようになっています。

> `(140,145]`

は「140を超えて145以下」となりますので、次の

> `(145,150]`

は同様に「145を超えて150以下」となりますので、階級間の境界値である`145`は`(140, 145]`側に入ります。つまり

* `(`は「超えて」、`)`は未満」（境界値を含まない）
* `[`は「以上」、`]`は「以下」（境界値を含む）

となっています。`cut()`関数の`include.lowest`オプションと`right`オプションを指定すると下表のように計算結果を変えることができます。

option                       | `right = TRUE`          | `right = FALSE` 
-----------------------------|-------------------------|-------------------
**`include.lowest = TRUE`**  | **`[`**`l,u] ... (l,u]` | `[l,u) ... [l,u]`
**`include.lowest = FALSE`** | **`(`**`l,u] ... (l,u]` | `[l,u) ... [l,u)`

　例えば最大値と最小値を境界値として含む階級を指定した場合、`include.lowest = FALSE, right = FALSE`と指定すると最大値が階級に含まれなくなります。
```{r}
with(x, cut(height, breaks = c(143, 155, 169),
            include.lowest = FALSE, right = FALSE))
```

　逆に`include.lowest = FALSE, right = TRUE`と指定すると最小値が階級に含まれなくなります。
```{r}
with(x, cut(height, breaks = c(143, 155, 169),
            include.lowest = FALSE, right = TRUE))
```

### 度数を求める
　`dplyr::count()`関数は`table()`関数と同様、個数を数える関数です。`table()`関数の計算対象はベクトル型ですが、`dplyr::count()`関数はデータフレーム型が計算対象になります。
```{r}
with(x, table(cut(height, breaks = pretty(height, n = nclass.Sturges(height)))))
```

```{r}
x %>% 
  # 階級を求めて、データを階級ごとに分ける
  dplyr::mutate(class = cut(height,
                            breaks = pretty(height, n = nclass.Sturges(height)))) %>% 
  # 階級ごとの度数を求める
  dplyr::count(class)
```

### 累積度数、相対度数、累積相対度数を求める
　`cumsum()`関数は累積和を計算する関数です。

```{r}
with(x, cumsum(table(cut(height, breaks = pretty(height, n = nclass.Sturges(height))))))
```

　`prop.table()`関数は相対度数を計算する関数です。
```{r}
with(x, prop.table(table(cut(height, breaks = pretty(height, n = nclass.Sturges(height))))))
```

　`prop.table()`関数の計算結果を`cumsum()`関数に渡すことで相対度数の累積和（累積相対度数）を求めることができます。
```{r}
with(x, cumsum(prop.table(table(cut(height, breaks = pretty(height, n = nclass.Sturges(height)))))))
```


### 階級値を求める
　階級値は階級の文字列から下限と上限の境界値を抜き出し、単純に下式で平均値を求めています。

$$\frac{(\mbox{下限} + 1) + \mbox{上限}}{2}$$

　手順としては以下のようにしています。

1. `as.character()`関数で 階級（`class`）を文字型に変換し、新しい変数（`class_value`）を作成する  
1. `tidyr::separate()`関数で文字型の変数（`class_value`）を`,`の前後で二つの変数（`l`, `h`）に分割する  
1. `stringr::str_remove`関数で二つの変数（`l`, `h`）から`(`や`]`を取り除き`as.numeric()`関数で文字列から数値に変換する  
1. 数値に変換された二つの変数から（`l`, `h`）平均値を求める  

```{r}
x %>% 
  # 階級を求めて、データを階級ごとに分ける
  dplyr::mutate(class = cut(height,
                            breaks = pretty(height, n = nclass.Sturges(height)),
                            include.lowest = FALSE, right = TRUE)) %>% 
  # 階級ごとの度数を求める
  dplyr::count(class) %>% 
  # 階級値を求める
  dplyr::mutate(class_value = as.character(class)) %>% 
  tidyr::separate(class_value, into = c("l", "h"), sep = ",") %>% 
  dplyr::mutate(l = as.numeric(stringr::str_remove(l, "[[:punct:]]")),
                h = as.numeric(stringr::str_remove(h, "[[:punct:]]"))) %>% 
  dplyr::mutate(mids = ((l + 1) + h) / 2)
```

The Tufte handout style is a style that Edward Tufte uses in his books and handouts. Tufte's style is known for its extensive use of sidenotes, tight integration of graphics with text, and well-set typography. This style has been implemented in LaTeX and HTML/CSS[^1], respectively. We have ported both implementations into the [**tufte** package](https://github.com/rstudio/tufte). If you want LaTeX/PDF output, you may use the `tufte_handout` format for handouts, and `tufte_book` for books. For HTML output, use `tufte_html`. These formats can be either specified in the YAML metadata at the beginning of an R Markdown document (see an example below), or passed to the `rmarkdown::render()` function. See @R-rmarkdown for more information about **rmarkdown**.

[^1]: See Github repositories [tufte-latex](https://github.com/tufte-latex/tufte-latex) and [tufte-css](https://github.com/edwardtufte/tufte-css)

``` yaml
---
title: "An Example Using the Tufte Style"
author: "John Smith"
output:
  tufte::tufte_handout: default
  tufte::tufte_html: default
---
```

There are two goals of this package:

1.  To produce both PDF and HTML output with similar styles from the same R Markdown document;
2.  To provide simple syntax to write elements of the Tufte style such as side notes and margin figures, e.g. when you want a margin figure, all you need to do is the chunk option `fig.margin = TRUE`, and we will take care of the details for you, so you never need to think about `\begin{marginfigure} \end{marginfigure}` or `<span class="marginfigure"> </span>`; the LaTeX and HTML code under the hood may be complicated, but you never need to learn or write such code.

If you have any feature requests or find bugs in **tufte**, please do not hesitate to file them to <https://github.com/rstudio/tufte/issues>. For general questions, you may ask them on StackOverflow: <https://stackoverflow.com/tags/rmarkdown>.

# Headings

This style provides first and second-level headings (that is, `#` and `##`), demonstrated in the next section. You may get unexpected output if you try to use `###` and smaller headings.

`r newthought('In his later books')`[^2], Tufte starts each section with a bit of vertical space, a non-indented paragraph, and sets the first few words of the sentence in small caps. To accomplish this using this style, call the `newthought()` function in **tufte** in an *inline R expression* `` `r ` `` as demonstrated at the beginning of this paragraph.[^3]

[^2]: [Beautiful Evidence](https://www.edwardtufte.com/tufte/books_be)

[^3]: Note you should not assume **tufte** has been attached to your R session. You should either `library(tufte)` in your R Markdown document before you call `newthought()`, or use `tufte::newthought()`.

# Figures

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures in the margin you can use the **knitr** chunk option `fig.margin = TRUE`. For example:

```{r fig-margin, fig.margin = TRUE, fig.cap = "MPG vs horsepower, colored by transmission.", fig.width=3.5, fig.height=3.5, cache=TRUE, message=FALSE}
library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() + geom_smooth() +
  theme(legend.position = 'bottom')
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can adjust the proportions of figures using the `fig.width` and `fig.height` chunk options. These are specified in inches, and will be automatically scaled down to fit within the handout margin.

## Arbitrary Margin Content

In fact, you can include anything in the margin using the **knitr** engine named `marginfigure`. Unlike R code chunks ```` ```{r} ````, you write a chunk starting with ```` ```{marginfigure} ```` instead, then put the content in the chunk. See an example on the right about the first fundamental theorem of calculus.

```{marginfigure}
We know from _the first fundamental theorem of calculus_ that for $x$ in $[a, b]$:
$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
```

For the sake of portability between LaTeX and HTML, you should keep the margin content as simple as possible (syntax-wise) in the `marginefigure` blocks. You may use simple Markdown syntax like `**bold**` and `_italic_` text, but please refrain from using footnotes, citations, or block-level elements (e.g. blockquotes and lists) there.

Note: if you set `echo = FALSE` in your global chunk options, you will have to add `echo = TRUE` to the chunk to display a margin figure, for example ```` ```{marginfigure, echo = TRUE} ````.

## Full Width Figures

You can arrange for figures to span across the entire page by using the chunk option `fig.fullwidth = TRUE`.

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "A full width figure.", warning=FALSE, message=FALSE, cache=TRUE}
ggplot(diamonds, aes(carat, price)) + geom_smooth() +
  facet_grid(~ cut)
```

Other chunk options related to figures can still be used, such as `fig.width`, `fig.cap`, `out.width`, and so on. For full width figures, usually `fig.width` is large and `fig.height` is small. In the above example, the plot size is $10 \times 2$.

## Arbitrary Full Width Content

Any content can span to the full width of the page. This feature requires Pandoc 2.0 or above. All you need is to put your content in a fenced `Div` with the class `fullwidth`, e.g.,

``` md
::: {.fullwidth}
Any _full width_ content here.
:::
```

Below is an example:

::: fullwidth
*R is free software and comes with ABSOLUTELY NO WARRANTY.* You are welcome to redistribute it under the terms of the GNU General Public License versions 2 or 3. For more information about these matters see <https://www.gnu.org/licenses/>.
:::

## Main Column Figures

Besides margin and full width figures, you can of course also include figures constrained to the main column. This is the default type of figures in the LaTeX/HTML output.

```{r fig-main, fig.cap = "A figure in the main column.", cache=TRUE}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```

# Sidenotes

One of the most prominent and distinctive features of this style is the extensive use of sidenotes. There is a wide margin to provide ample room for sidenotes and small figures. Any use of a footnote will automatically be converted to a sidenote. [^4]

[^4]: This is a sidenote that was entered using a footnote.

If you'd like to place ancillary information in the margin without the sidenote mark (the superscript number), you can use the `margin_note()` function from **tufte** in an inline R expression. `r margin_note("This is a margin note.  Notice that there is no number preceding the note.")` This function does not process the text with Pandoc, so Markdown syntax will not work here. If you need to write anything in Markdown syntax, please use the `marginfigure` block described previously.

# References

References can be displayed as margin notes for HTML output. For example, we can cite R here [@R-base]. To enable this feature, you must set `link-citations: yes` in the YAML metadata, and the version of `pandoc-citeproc` should be at least 0.7.2. You can always install your own version of Pandoc from <https://pandoc.org/installing.html> if the version is not sufficient. To check the version of `pandoc-citeproc` in your system, you may run this in R:

```{r eval=FALSE}
system2('pandoc-citeproc', '--version')
```

If your version of `pandoc-citeproc` is too low, or you did not set `link-citations: yes` in YAML, references in the HTML output will be placed at the end of the output document.

# Tables

You can use the `kable()` function from the **knitr** package to format tables that integrate well with the rest of the Tufte handout style. The table captions are placed in the margin like figures in the HTML output.

```{r}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```

# Block Quotes

We know from the Markdown syntax that paragraphs that start with `>` are converted to block quotes. If you want to add a right-aligned footer for the quote, you may use the function `quote_footer()` from **tufte** in an inline R expression. Here is an example:

> "If it weren't for my lawyer, I'd still be in prison. It went a lot faster with two people digging."
>
> `r tufte::quote_footer('--- Joe Martin')`

Without using `quote_footer()`, it looks like this (the second line is just a normal paragraph):

> "Great people talk about ideas, average people talk about things, and small people talk about wine."
>
> --- Fran Lebowitz

# Responsiveness

The HTML page is responsive in the sense that when the page width is smaller than 760px, sidenotes and margin notes will be hidden by default. For sidenotes, you can click their numbers (the superscripts) to toggle their visibility. For margin notes, you may click the circled plus signs to toggle visibility.

# More Examples

The rest of this document consists of a few test cases to make sure everything still works well in slightly more complicated scenarios. First we generate two plots in one figure environment with the chunk option `fig.show = 'hold'`:

```{r fig-two-together, fig.cap="Two plots in one figure environment.", fig.show='hold', cache=TRUE, message=FALSE}
p <- ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point()
p
p + geom_smooth()
```

Then two plots in separate figure environments (the code is identical to the previous code chunk, but the chunk option is the default `fig.show = 'asis'` now):

```{r fig-two-separate, ref.label='fig-two-together', fig.cap=sprintf("Two plots in separate figure environments (the %s plot).", c("first", "second")), cache=TRUE, message=FALSE}
```

You may have noticed that the two figures have different captions, and that is because we used a character vector of length 2 for the chunk option `fig.cap` (something like `fig.cap = c('first plot', 'second plot')`).

Next we show multiple plots in margin figures. Similarly, two plots in the same figure environment in the margin:

```{r fig-margin-together, fig.margin=TRUE, fig.show='hold', fig.cap="Two plots in one figure environment in the margin.", fig.width=3.5, fig.height=2.5, cache=TRUE}
p
p + geom_smooth(method = 'lm')
```

Then two plots from the same code chunk placed in different figure environments:

```{r fig-margin-separate, fig.margin=TRUE, fig.cap=sprintf("Two plots in separate figure environments in the margin (the %s plot).", c("first", "second")), fig.width=3.5, fig.height=2.5, cache=TRUE}
knitr::kable(head(iris, 15))
p
knitr::kable(head(iris, 12))
p + geom_smooth(method = 'lm')
knitr::kable(head(iris, 5))
```

We blended some tables in the above code chunk only as *placeholders* to make sure there is enough vertical space among the margin figures, otherwise they will be stacked tightly together. For a practical document, you should not insert too many margin figures consecutively and make the margin crowded.

You do not have to assign captions to figures. We show three figures with no captions below in the margin, in the main column, and in full width, respectively.

```{r fig-nocap-margin, fig.margin=TRUE, fig.width=3.5, fig.height=2, cache=TRUE}
# a boxplot of weight vs transmission; this figure
# will be placed in the margin
ggplot(mtcars2, aes(am, wt)) + geom_boxplot() +
  coord_flip()
```

```{r fig-nocap-main, cache=TRUE}
# a figure in the main column
p <- ggplot(mtcars, aes(wt, hp)) + geom_point()
p
```

```{r fig-nocap-fullwidth, fig.fullwidth=TRUE, fig.width=10, fig.height=3, cache=TRUE}
# a fullwidth figure
p + geom_smooth(method = 'lm') + facet_grid(~ gear)
```

# Some Notes on Tufte CSS

There are a few other things in Tufte CSS that we have not mentioned so far. If you prefer `r sans_serif('sans-serif fonts')`, use the function `sans_serif()` in **tufte**. For epigraphs, you may use a pair of underscores to make the paragraph italic in a block quote, e.g.

> *I can win an argument on any topic, against any opponent. People know this, and steer clear of me at parties. Often, as a sign of their great respect, they don't even invite me.*
>
> `r quote_footer('--- Dave Barry')`

We hope you will enjoy the simplicity of R Markdown and this R package, and we sincerely thank the authors of the Tufte-CSS and Tufte-LaTeX projects for developing the beautiful CSS and LaTeX classes. Our **tufte** package would not have been possible without their heavy lifting.

You can turn on/off some features of the Tufte style in HTML output. The default features enabled are:

``` yaml
output:
  tufte::tufte_html:
    tufte_features: ["fonts", "background", "italics"]
```

If you do not want the page background to be lightyellow, you can remove `background` from `tufte_features`. You can also customize the style of the HTML page via a CSS file. For example, if you do not want the subtitle to be italic, you can define

``` css
h3.subtitle em {
  font-style: normal;
}
```

in, say, a CSS file `my_style.css` (under the same directory of your Rmd document), and apply it to your HTML output via the `css` option, e.g.,

``` yaml
output:
  tufte::tufte_html:
    tufte_features: ["fonts", "background"]
    css: "my_style.css"
```

There is also a variant of the Tufte style in HTML/CSS named "[Envisoned CSS](https://github.com/nogginfuel/envisioned-css)". This style can be used by specifying the argument `tufte_variant = 'envisioned'` in `tufte_html()`[^5], e.g.

[^5]: The actual Envisioned CSS was not used in the **tufte** package. We only changed the fonts, background color, and text color based on the default Tufte style.

``` yaml
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
```

To see the R Markdown source of this example document, you may follow [this link to Github](https://github.com/rstudio/tufte/raw/main/inst/rmarkdown/templates/tufte_html/skeleton/skeleton.Rmd), use the wizard in RStudio IDE (`File -> New File -> R Markdown -> From Template`), or open the Rmd file in the package:

```{r eval=FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

This document is also available in [Chinese](https://rstudio.github.io/tufte/cn/), and its `envisioned` style can be found [here](https://rstudio.github.io/tufte/envisioned/).

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```

<!--chapter:end:section_01.Rmd-->

